### 2022-12-05 dissertation/chapters/08-results.tex
✨ (review): Spelling and grammar, derived publications fix

Although the different configurations of the \acs{BHH} are all able to succesfully train the underlying \acs{FFNN}, in this particular case, there is no clear distinction between the performance of any of the configurations under consideration.

@@

The first logical observation to be made from Figures~\ref{fig:results:case_study:alphas:0} to \ref{fig:results:case_study:alphas:8} is the step-wise, increasing nature of $\boldsymbol{\alpha}$ for the bhh\_replay\_250 configuration, illustrated in \textcolor{green}{green}. Since the replay window size is sufficienty large to contain the performance log for all the steps executed in the training process, the \acs{BHH} does not forget past performances of low-level heuristics at all. The value for $\boldsymbol{\alpha}$ is never reset to its initial value of 1.0, and thus, $\boldsymbol{\alpha}$ continues to increase throughout the training process. However, it should be noted that the rate and degree of change is different for different indices of $\boldsymbol{\alpha}$. The aforementioned observation is the first indicator of the learning process yielded by the \acs{BHH}. Heuristics that perform well will see their corresponding element in $\boldsymbol{\alpha}$ increase more rapidly than other heuristics that do not perform well.

The next observations that can be made are for the bhh\_baseline configuration, illustrated in \textcolor{red}{red}, and the bhh\_credit\_symmetric configuration in \textcolor{blue}{blue}. Both these configurations see $\boldsymbol{\alpha}$ being reset to its initial value of 1.0 as regular intervals. This interval is defined by the \textit{reanalysis interval} hyper-parameter. The reanalysis interval dictates the frequency at which Bayesian analysis is conducted on the performance log, mainted by the \acs{BHH}. Bayesian analysis is used to update $\boldsymbol{\alpha}$ only at the reanalysis interval. As a result, small plateaus appear where $\boldsymbol{\alpha}$ does not change.

@@

Furthermore, it is important to mention that a change in $\boldsymbol{\alpha}$ between reanalysis windows does not yet necessarily indicate that the \acs{BHH} is learning. The concentration parameter $\boldsymbol{\alpha}$ tracks the \textit{occurrence} of the random event of observing heuristics (denoted $\boldsymbol{{H}}$). \index{heuristic}Heuristics can be observed as a result of good performance, by which the \acs{BHH} then learns to frequenctly reselect these \index{heuristic}heuristics again, or just by chance through the stochastic nature of probabilistic sampling as implemented by the \acs{BHH}. Further investigation is required to illustrate the learning mechanism of the \acs{BHH}.

@@

This section provides a detailed investigation into the \textit{probability distribution of \index{heuristic}heuristic selection probabilities} as it changes throughout the training process. As a reminder, the \acs{BHH} implements a Bayesian view of probabilistic modeling and thus, \index{heuristic}heuristic selection probabilities are defined by an underlying probablistic distribution themselves. The probability distribution of \index{heuristic}heuristic selection probabilities is denoted by $P(\boldsymbol{\theta} \vert \boldsymbol{\alpha})$, where $\boldsymbol{\theta} \sim Dir(\boldsymbol{\alpha, K})$ are the sampled heuristic selection probabilities.

@@

From the illustrations presented in Figures~\ref{fig:results:case_study:thetas:0} to \ref{fig:results:case_study:thetas:8}, a clearer picture of the learning process of the \acs{BHH} is formed. The first important observation to make is for the bhh\_replay\_250 configuration, presented in \textcolor{green}{green}. As a reminder, ten low-level \index{heuristic}heuristics are included in the heuristic pool, yielding an expected mean \index{heuristic}heuristics selection probability of 0.1, for each \index{heuristic}heuristic, by the frequentist view of probabilistic modeling.

@@

Figures~\ref{fig:results:case_study:p_HgEC:0:0} to \ref{fig:results:case_study:p_HgEC:0:8} provide illustrations of the calculated posterior \index{heuristic}heuristic selection probabilities at indices 0, 6, 7, and 8, throughout the training process, averaged over 30 runs. Similar to before, these illustrations are also presented in log scale. As before, illustrations for the other indices are left out for brevity as they yield simular illustrations.

@@

The main observation to make from Figures~\ref{fig:results:case_study:p_HgEC:0:0} to \ref{fig:results:case_study:p_HgEC:0:8} is that the implemented posterior heuristic selection distribution, defined by $P(\boldsymbol{H} \vert \boldsymbol{E}, \boldsymbol{C}; \boldsymbol{\theta}, \boldsymbol{\phi}, \boldsymbol{\psi})$, does not yield normalised probabilities, but rather yield unnormalised \textit{logits}, which are used to parameterise a \index{Categorical probability distribution}Categorical probability distribution from which new \index{heuristic}heuristic selections are sampled. The reasons for the aforementioned is because the probablistic model is evaluated proportionaly as was discussed in Chapter~\ref{chap:bhh}. As a reminder, the log-sum-exp trick is used in order to maintain numerical stability, yielding \textit{logits} instead of probabilities.

Another observation to make is for the bhh\_replay\_250 configuration $(\textcolor{green}{green})$. Figures~\ref{fig:results:case_study:p_HgEC:0:0} to \ref{fig:results:case_study:p_HgEC:0:8} show that the posterior heuristic selection probabilities converge to the expected heuristic selection probabilities later in the training stages. The aforementioned suggests that the \acs{BHH} is not able to find anymore performance biases and cannot exploit better solutions. At this point, the \acs{BHH} starts to explore more as the concentration parameters are reanalysed more uniformly, resolving more and more to a random search in attempt to find better solutions.

@@

Table~\ref{tab:results:standalone:metrics:rank} provides the empirical results from Table~\ref{tab:results:standalone:metrics:loss} in ranked format. The performance rank is calculated as the average rank produced by each \index{heuristic}heuristic, across all datasets, for all indepented runs and all epochs. The average rank across all epochs produces a view on the performance of the \index{heuristic}heuristics as it relates to the entire training process. Finally, a normalised average rank is provided for the overall performance of all \index{heuristic}heuristics at the bottom of the table. The normalised average rank is calculated as a discrete normalisation of the average rank achieved across all datasets, for all independent runs and epochs.

@@

From the normalised average ranks provided in Table~\ref{tab:results:standalone:metrics:rank}, it can be seen that the bhh\_gd configuration ranked fourth, while the bhh\_all and bhh\_mh configurations ranked sixth and eigth amongst all thirteen \index{heuristic}heuristic implementations respectively. These results show that the \acs{BHH} generally performs well, but is not able to outperform the best heuristic for each dataset.

@@

From Figures~\ref{fig:results:heuristic_pool:descriptive:descriptive} and~\ref{fig:results:heuristic_pool:descriptive:cd}, it is clear that the \textit{gd} configuration configuration performs best overall, with exceptions to the parkinsons and housing datasets, where the \textit{all} and \textit{gd} configurations performed equally well with statistically insignificant differences in their outcomes.

@@

From Table \ref{tab:results:population:metrics:rank} it can be seen that a lower population size of five, mostly produced the best results, with exception to the bike and car datasets, for which the lowest population size configuration produced the worst performance, yielding statistically significant differences in outcomes from the other datasets. A larger population size configuration is prefered for the bike and car datasets. Furthermore, a population size configuration of ten, slightly higher than the default of five, is preferred for the forest fires and housing datasets. Finally, the overall normalised average rank is provided at the bottom of the table, showing that a population size of five produced the best performance outcome across all datasets on average.

@@

Divergence of the training loss can be observed in Figure \ref{fig:results:population:figures:loss:test:mushroom}. Note that, despite the divergence of training loss, the accuracy is still almost 100\% and that these divergent behaviours are simply a result of an attempt to further improve performance, by exploring the heuristic space more. For example, heuristics could have momentum as a result of good training progression thus far, but then overshoot good solutions after \index{heuristic}heuristic reselection. Overshooting good solutions then causes the \index{heuristic}heuristics to struggle to converge back to good solutions, especially if many different \index{heuristic}heuristics are selected in quick succession. A suggestion to this problem is to implement a move-acceptance strategy as was mentioned before. At the nineth epoch, the \acs{BHH} fails to produce better solutions on the training set. Since there is not much room for more improvement, the \acs{BHH} is bound to try different \index{heuristic}heuristics that yield sub-optimal results. From this point onwards, the \acs{BHH} finds slightly better results on the train set, and as a result, produces volatile performance on the test set.

@@

To illustrate the effects of the reselection interval hyper-paramter on the entire trianing, process, Table \ref{tab:results:reselection:metrics:rank} provides the average ranked performance, per dataset, of each of the \acs{BHH} reselection interval configurations. Similar to before, the performance rank is calculated as the average rank produced by reselection interval configuration, for all datasets, over all independent runs and epochs.

@@

Furthermore, the smaller the reanalysis interval, the more frequently the \acs{BHH} resets concentration parameters. The smaller the replay window size, the less performance samples in the performance log are considered for reanalysis. A replay window size of one yields a case where only the last executed step is considered for reanalysis and thus provides a shortsighted view on the relevance of past performance. As such, reselection is based only on the most recent performance of each selected \index{heuristic}heuristic. However, if the reanalysis interval is larger than the replay window size, the \acs{BHH} does not update its beliefs fast enough, and as a result, \index{heuristic}heuristic performance evidence is lost.

@@

Figure~\ref{fig:results:reanalysis:figures:parkinsons} provides the train and test loss plots for an example regression dataset (parkinsons) as it relates to the renalysis interval experimental group. As before, the illustrations are provided in log scale and illustrations of the train and test loss plots for the other regression datasets are left out for brevity.

>>>>

Although the different configurations of the \acs{BHH} are all able to successfully train the underlying \acs{FFNN}, in this particular case, there is no clear distinction between the performance of any of the configurations under consideration.

@@

The first logical observation to be made from Figures~\ref{fig:results:case_study:alphas:0} to \ref{fig:results:case_study:alphas:8} is the step-wise, increasing nature of $\boldsymbol{\alpha}$ for the bhh\_replay\_250 configuration, illustrated in \textcolor{green}{green}. Since the replay window size is sufficiently large to contain the performance log for all the steps executed in the training process, the \acs{BHH} does not forget past performances of low-level heuristics at all. The value for $\boldsymbol{\alpha}$ is never reset to its initial value of 1.0, and thus, $\boldsymbol{\alpha}$ continues to increase throughout the training process. However, it should be noted that the rate and degree of change is different for different indices of $\boldsymbol{\alpha}$. The aforementioned observation is the first indicator of the learning process yielded by the \acs{BHH}. Heuristics that perform well will see their corresponding element in $\boldsymbol{\alpha}$ increase more rapidly than other heuristics that do not perform well.

The next observations that can be made are for the bhh\_baseline configuration, illustrated in \textcolor{red}{red}, and the bhh\_credit\_symmetric configuration in \textcolor{blue}{blue}. Both these configurations see $\boldsymbol{\alpha}$ being reset to its initial value of 1.0 as regular intervals. This interval is defined by the \textit{reanalysis interval} hyper-parameter. The reanalysis interval dictates the frequency at which Bayesian analysis is conducted on the performance log, maintained by the \acs{BHH}. Bayesian analysis is used to update $\boldsymbol{\alpha}$ only at the reanalysis interval. As a result, small plateaus appear where $\boldsymbol{\alpha}$ does not change.

@@

Furthermore, it is important to mention that a change in $\boldsymbol{\alpha}$ between reanalysis windows does not yet necessarily indicate that the \acs{BHH} is learning. The concentration parameter $\boldsymbol{\alpha}$ tracks the \textit{occurrence} of the random event of observing heuristics (denoted $\boldsymbol{{H}}$). \index{heuristic}Heuristics can be observed as a result of good performance, by which the \acs{BHH} then learns to frequently reselect these \index{heuristic}heuristics again, or just by chance through the stochastic nature of probabilistic sampling as implemented by the \acs{BHH}. Further investigation is required to illustrate the learning mechanism of the \acs{BHH}.

@@

This section provides a detailed investigation into the \textit{probability distribution of \index{heuristic}heuristic selection probabilities} as it changes throughout the training process. As a reminder, the \acs{BHH} implements a Bayesian view of probabilistic modelling and thus, \index{heuristic}heuristic selection probabilities are defined by an underlying probabilistic distribution themselves. The probability distribution of \index{heuristic}heuristic selection probabilities is denoted by $P(\boldsymbol{\theta} \vert \boldsymbol{\alpha})$, where $\boldsymbol{\theta} \sim Dir(\boldsymbol{\alpha, K})$ are the sampled heuristic selection probabilities.

@@

From the illustrations presented in Figures~\ref{fig:results:case_study:thetas:0} to \ref{fig:results:case_study:thetas:8}, a clearer picture of the learning process of the \acs{BHH} is formed. The first important observation to make is for the bhh\_replay\_250 configuration, presented in \textcolor{green}{green}. As a reminder, ten low-level \index{heuristic}heuristics are included in the heuristic pool, yielding an expected mean \index{heuristic}heuristics selection probability of 0.1, for each \index{heuristic}heuristic, by the frequentist view of probabilistic modelling.

@@

Figures~\ref{fig:results:case_study:p_HgEC:0:0} to \ref{fig:results:case_study:p_HgEC:0:8} provide illustrations of the calculated posterior \index{heuristic}heuristic selection probabilities at indices 0, 6, 7, and 8, throughout the training process, averaged over 30 runs. Similar to before, these illustrations are also presented in log scale. As before, illustrations for the other indices are left out for brevity as they yield similar illustrations.

@@

The main observation to make from Figures~\ref{fig:results:case_study:p_HgEC:0:0} to \ref{fig:results:case_study:p_HgEC:0:8} is that the implemented posterior heuristic selection distribution, defined by $P(\boldsymbol{H} \vert \boldsymbol{E}, \boldsymbol{C}; \boldsymbol{\theta}, \boldsymbol{\phi}, \boldsymbol{\psi})$, does not yield normalised probabilities, but rather yield unnormalised \textit{logits}, which are used to parameterise a \index{Categorical probability distribution}Categorical probability distribution from which new \index{heuristic}heuristic selections are sampled. The reasons for the aforementioned is because the probabilistic model is evaluated proportionally as was discussed in Chapter~\ref{chap:bhh}. As a reminder, the log-sum-exp trick is used in order to maintain numerical stability, yielding \textit{logits} instead of probabilities.

Another observation to make is for the bhh\_replay\_250 configuration $(\textcolor{green}{green})$. Figures~\ref{fig:results:case_study:p_HgEC:0:0} to \ref{fig:results:case_study:p_HgEC:0:8} show that the posterior heuristic selection probabilities converge to the expected heuristic selection probabilities later in the training stages. The aforementioned suggests that the \acs{BHH} is not able to find further performance biases and cannot exploit better solutions. At this point, the \acs{BHH} starts to explore more as the concentration parameters are reanalysed more uniformly, resolving more and more to a random search in attempt to find better solutions.

@@

Table~\ref{tab:results:standalone:metrics:rank} provides the empirical results from Table~\ref{tab:results:standalone:metrics:loss} in ranked format. The performance rank is calculated as the average rank produced by each \index{heuristic}heuristic, across all datasets, for all independent runs and all epochs. The average rank across all epochs produces a view on the performance of the \index{heuristic}heuristics as it relates to the entire training process. Finally, a normalised average rank is provided for the overall performance of all \index{heuristic}heuristics at the bottom of the table. The normalised average rank is calculated as a discrete normalisation of the average rank achieved across all datasets, for all independent runs and epochs.

@@

From the normalised average ranks provided in Table~\ref{tab:results:standalone:metrics:rank}, it can be seen that the bhh\_gd configuration ranked fourth, while the bhh\_all and bhh\_mh configurations ranked sixth and eighth amongst all thirteen \index{heuristic}heuristic implementations respectively. These results show that the \acs{BHH} generally performs well, but is not able to outperform the best heuristic for each dataset.

@@

From Figures~\ref{fig:results:heuristic_pool:descriptive:descriptive} and~\ref{fig:results:heuristic_pool:descriptive:cd}, it is clear that the \textit{gd} configuration performs best overall, with exceptions to the parkinsons and housing datasets, where the \textit{all} and \textit{gd} configurations performed equally well with statistically insignificant differences in their outcomes.

@@

From Table \ref{tab:results:population:metrics:rank} it can be seen that a lower population size of five, mostly produced the best results, with exception to the bike and car datasets, for which the lowest population size configuration produced the worst performance, yielding statistically significant differences in outcomes from the other datasets. A larger population size configuration is preferred for the bike and car datasets. Furthermore, a population size configuration of ten, slightly higher than the default of five, is preferred for the forest fires and housing datasets. Finally, the overall normalised average rank is provided at the bottom of the table, showing that a population size of five produced the best performance outcome across all datasets on average.

@@

Divergence of the training loss can be observed in Figure \ref{fig:results:population:figures:loss:test:mushroom}. Note that, despite the divergence of training loss, the accuracy is still almost 100\% and that these divergent behaviours are simply a result of an attempt to further improve performance, by exploring the heuristic space more. For example, heuristics could have momentum as a result of good training progression thus far, but then overshoot good solutions after \index{heuristic}heuristic reselection. Overshooting good solutions then causes the \index{heuristic}heuristics to struggle to converge back to good solutions, especially if many different \index{heuristic}heuristics are selected in quick succession. A suggestion to this problem is to implement a move-acceptance strategy as was mentioned before. At the ninth epoch, the \acs{BHH} fails to produce better solutions on the training set. Since there is not much room for more improvement, the \acs{BHH} is bound to try different \index{heuristic}heuristics that yield sub-optimal results. From this point onwards, the \acs{BHH} finds slightly better results on the train set, and as a result, produces volatile performance on the test set.

@@

To illustrate the effects of the reselection interval hyper-parameter on the entire training, process, Table \ref{tab:results:reselection:metrics:rank} provides the average ranked performance, per dataset, of each of the \acs{BHH} reselection interval configurations. Similar to before, the performance rank is calculated as the average rank produced by reselection interval configuration, for all datasets, over all independent runs and epochs.

@@

Furthermore, the smaller the reanalysis interval, the more frequently the \acs{BHH} resets concentration parameters. The smaller the replay window size, the less performance samples in the performance log are considered for reanalysis. A replay window size of one yields a case where only the last executed step is considered for reanalysis and thus provides a short-sighted view on the relevance of past performance. As such, reselection is based only on the most recent performance of each selected \index{heuristic}heuristic. However, if the reanalysis interval is larger than the replay window size, the \acs{BHH} does not update its beliefs fast enough, and as a result, \index{heuristic}heuristic performance evidence is lost.

@@

Figure~\ref{fig:results:reanalysis:figures:parkinsons} provides the train and test loss plots for an example regression dataset (parkinsons) as it relates to the reanalysis interval experimental group. As before, the illustrations are provided in log scale and illustrations of the train and test loss plots for the other regression datasets are left out for brevity.

### 2022-12-05 dissertation/chapters/09-conclusion.tex
✨ (review): Spelling and grammar, derived publications fix

The research presented in this disseration identified the possibility of using a different approach, referred to as \acp{HH}, to automate the \index{heuristic}heuristic selection process. \acp{HH} are high-level search algorithms that search in the \index{heuristic}heuristic space, whereby low-level \index{heuristic}heuristics search in the solution space. \acp{HH} have been shown to provide a general solution to the \index{heuristic}heuristic selection process. The benefit that such an approach provides is that it automates the \index{heuristic}heuristic selection process. This approach saves researchers time and potentially produce a solution that could generalise to multiple problems.

@@

      \item An empirical process was desiged and the detail of the implementation for the empirical process and the \acs{BHH} was provided.

@@

All experimental groups were run for a maximum of 30 epochs in order to allow for the studying of the \ac{BHH} even after training has converged and overfitting has occured. The maximum epoch strategy yielded a training process that is longer than what was empirically found to be necessary. This also meant that no early stopping strategy was used. Future research opportunities should incorporate an early stopping strategy of the training process.

@@

Various illustrations of the train and test loss plots, as it relates to training of a \acs{FFNN} on the the iris dataset, have been provided. Furthermore, various illustrations on the changes of the \acs{BHH} state parameters were provided for visual aid.

The main findings that were made for this experimental group are given as follows: Most of the training progression is made in the early stages of the training process. That leads to the conclusion that the \acs{BHH} has a small window from which it should learn and gain the most in the training process. After training has converged, the \acs{BHH} resets its concentration parameters and heuristic selection returns to the symmetric heuristic selection case. As such, the \acs{BHH} explores other heuristics in an attempt to further improve on the current best solution found. The test set was used as a validation set during training. Some overfitting can be observed as the \acs{BHH} tried finding better solutions on the train set, but at the cost of generalisation on the test set. Minor divergence of the training loss is observed as the \acs{BHH} explores other heuristics in an attempt to improve performance. Since no move-acceptance strategy, and no early stopping stategy was used, the \acs{BHH} could select heuristics that are sub-optimal. Future research opportunities should incorporate these aforementioned strategies.

@@

For this experimental group, the same three \index{heuristic pool}heuristical pool variants as outlined for the \acs{BHH} behavioural case study, was implemented. These variants are denoted as \textit{all}, \textit{gd}, and \textit{mh}. This experimental group then provided an opportunity to look at the effects of the \index{heuristic pool}heuristic pool hyper-parameter in more detail.

@@

For this experimental group, five different reselection intervals were evaluated. These included reselection intervals of 1, 5, 10, 15, and 20. Overall, with statistical significance, it was found that a larger reselection interval is generally preferred. It was found that larger reselection intervals lead to longer timeframes for the low-level heuristics to progress, smoothing out update steps. Larger reselection intervals also allow the \acs{BHH} to collect more samples of \index{heuristic}heuristic performance evidence from which it can learn. Therefore, a conclusion that can be made is that \index{heuristic}heuristic and solution exploitation is preferred over exploration. As before, illustrations of the train and test loss and accuracy plots for some example datasets were provided.

>>>>

The research presented in this dissertation identified the possibility of using a different approach, referred to as \acp{HH}, to automate the \index{heuristic}heuristic selection process. \acp{HH} are high-level search algorithms that search in the \index{heuristic}heuristic space, whereby low-level \index{heuristic}heuristics search in the solution space. \acp{HH} have been shown to provide a general solution to the \index{heuristic}heuristic selection process. The benefit that such an approach provides is that it automates the \index{heuristic}heuristic selection process. This approach saves researchers time and potentially produce a solution that could generalise to multiple problems.

@@

      \item An empirical process was designed and the detail of the implementation for the empirical process and the \acs{BHH} was provided.

@@

All experimental groups were run for a maximum of 30 epochs in order to allow for the studying of the \ac{BHH} even after training has converged and overfitting has occurred. The maximum epoch strategy yielded a training process that is longer than what was empirically found to be necessary. This also meant that no early stopping strategy was used. Future research opportunities should incorporate an early stopping strategy of the training process.

@@

Various illustrations of the train and test loss plots, as it relates to training of a \acs{FFNN} on the iris dataset, have been provided. Furthermore, various illustrations on the changes of the \acs{BHH} state parameters were provided for visual aid.

The main findings that were made for this experimental group are given as follows: Most of the training progression is made in the early stages of the training process. That leads to the conclusion that the \acs{BHH} has a small window from which it should learn and gain the most in the training process. After training has converged, the \acs{BHH} resets its concentration parameters and heuristic selection returns to the symmetric heuristic selection case. As such, the \acs{BHH} explores other heuristics in an attempt to further improve on the current best solution found. The test set was used as a validation set during training. Some overfitting can be observed as the \acs{BHH} tried finding better solutions on the train set, but at the cost of generalisation on the test set. Minor divergence of the training loss is observed as the \acs{BHH} explores other heuristics in an attempt to improve performance. Since no move-acceptance strategy, and no early stopping strategy was used, the \acs{BHH} could select heuristics that are sub-optimal. Future research opportunities should incorporate these aforementioned strategies.

@@

For this experimental group, the same three \index{heuristic pool}heuristic pool variants as outlined for the \acs{BHH} behavioural case study, was implemented. These variants are denoted as \textit{all}, \textit{gd}, and \textit{mh}. This experimental group then provided an opportunity to look at the effects of the \index{heuristic pool}heuristic pool hyper-parameter in more detail.

@@

For this experimental group, five different reselection intervals were evaluated. These included reselection intervals of 1, 5, 10, 15, and 20. Overall, with statistical significance, it was found that a larger reselection interval is generally preferred. It was found that larger reselection intervals lead to longer time frames for the low-level heuristics to progress, smoothing out update steps. Larger reselection intervals also allow the \acs{BHH} to collect more samples of \index{heuristic}heuristic performance evidence from which it can learn. Therefore, a conclusion that can be made is that \index{heuristic}heuristic and solution exploitation is preferred over exploration. As before, illustrations of the train and test loss and accuracy plots for some example datasets were provided.
